Open archaeology, open source?

### 2023-10-12 analysis/_03-data_methodology.qmd
Add statistics on lifespan and longevity

# Amalgamate categories into fewer rough categories
# TODO: redo
oarch <- mutate(oarch, rough_category = recode(category,
                                 "R" = "Packages (R, Python, etc.)",
                                 "Web apps" = "Standalone apps",
                                 "Datasets" = "Analyses & datasets",
                                 "Tutorials" = "Lists & tutorials",
                                 "Python" = "Packages (R, Python, etc.)",
                                 "Desktop apps" = "Standalone apps",
                                 "QGIS" = "Plugins",
                                 "Lists" = "Lists & tutorials",
                                 "Mobile apps" = "Standalone apps",
                                 "Analyses" = "Analyses & datasets",
                                 "Spreadsheets" = "Forms & spreadsheets",
                                 "Command line tools" = "Standalone apps",
                                 "Plugins" = "Plugins",
                                 "ArcGis" = "Plugins",
                                 "Filemaker" = "Plugins",
                                 "ODK" = "Forms & spreadsheets",
                                 "Lisp" = "Packages (R, Python, etc.)",
                                 .default = "Other",
                                 .missing = "Other"
                                 ))
```

>>>>

# Amalgamate categories into fewer rough categories
# TODO: redo
oarch <- mutate(oarch, rough_category = recode(category,
                                 "R" = "Packages (R, Python, etc.)",
                                 "Web apps" = "Standalone apps",
                                 "Datasets" = "Analyses & datasets",
                                 "Tutorials" = "Lists & tutorials",
                                 "Python" = "Packages (R, Python, etc.)",
                                 "Desktop apps" = "Standalone apps",
                                 "QGIS" = "Plugins",
                                 "Lists" = "Lists & tutorials",
                                 "Mobile apps" = "Standalone apps",
                                 "Analyses" = "Analyses & datasets",
                                 "Spreadsheets" = "Forms & spreadsheets",
                                 "Command line tools" = "Standalone apps",
                                 "Plugins" = "Plugins",
                                 "ArcGis" = "Plugins",
                                 "Filemaker" = "Plugins",
                                 "ODK" = "Forms & spreadsheets",
                                 "Lisp" = "Packages (R, Python, etc.)",
                                 .default = "Other",
                                 .missing = "Other"
                                 ))

# TODO: exclude peripheral repos that throw things off
# * Big boys like QField (see _06-network.qmd)
# * ArchaeoLines – ends up picking up the whole stellarium repo, which is huge
# * aDNA??
```

### 2023-10-12 analysis/_05-collaboration.qmd
Add statistics on lifespan and longevity

```{r generate-oarch-lifespan}
oarch %>% 
  unnest(gh_commits) %>% 
  drop_na(datetime) %>% 
  group_by(item_name) %>%
  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
  summarise(
    category = first(category),
    n_contribs = first(n_contribs),
    n = n(),
    first_commit = min(datetime),
    last_commit = max(datetime),
    lifespan = as.integer(last_commit - first_commit),
    commit_rate = n / lifespan
  ) ->
  oarch_lifespan
```

```{r fig-lifespan-hist}
#| fig-cap: Project lifespan 
oarch_lifespan %>%
  ggplot(aes(x = lifespan)) +
  geom_histogram(binwidth = 100) +
  scale_x_continuous(breaks = scales::breaks_width(1000)) +
  labs(
    x = "Lifespan (days)", y = "Repositories",
  ) +
  theme_oarch()
```

We measured the lifespan of a repository as the time between the first and latest commit, and its activity as the rate of commits.
Here therefore refer to the *development* lifespan of a project—measured here as the time between a repository's first and last commit—which is not necessary linked to its use-life.
The lifespan and activity of repositories in open-archaeo vary greatly (@fig-lifespan).
The average project lasts <!-- TODO --> days with <!-- TODO --> commits per day.
<!-- TODO: turn all these percentages into code -->
Many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
The longest are <!-- TODO -->.
Commit rates vary between <!-- TODO -->, with peaks at <!-- TODO -->.

>>>>

```{r generate-oarch-lifespan}
oarch %>% 
  unnest(gh_commits) %>% 
  drop_na(datetime) %>% 
  group_by(item_name) %>%
  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
  summarise(
    category = first(category),
    n_contribs = first(n_contribs),
    n = n(),
    first_commit = min(datetime),
    last_commit = max(datetime),
    lifespan = as.integer(last_commit - first_commit),
    commit_rate = n / lifespan
  ) |>
  mutate(commit_rate = if_else(is.finite(commit_rate), commit_rate, NA)) ->
  oarch_lifespan

# Statistics
n_1commit <- nrow(filter(oarch_lifespan, n == 1))

p_lifespan_30days <- nrow(filter(oarch_lifespan, lifespan < 30)) / nrow(oarch_lifespan)
p_lifespan_90days <- nrow(filter(oarch_lifespan, lifespan < 90)) / nrow(oarch_lifespan)
p_lifespan_365days <- nrow(filter(oarch_lifespan, lifespan < 365)) / nrow(oarch_lifespan)

oarch_lifespan |>
  arrange(-lifespan) |>
  slice(1:10) |>
  pull(lifespan) |>
  range() ->
  lifespan_top10_range

max_commit_rate <- max(oarch_lifespan$commit_rate, na.rm = TRUE)
p_commit_rate_1 <- nrow(filter(oarch_lifespan, commit_rate < 1)) / nrow(oarch_lifespan)
```

We measured the lifespan of a repository as the time between the first and latest commit, and its activity as the rate of commits.
Here therefore we refer to the *development* lifespan of a project, which is not necessary related to its use-life.
By these metrics, the lifespan and activity of repositories in open-archaeo vary greatly (@fig-lifespan).
The average project lasts `r round(mean(oarch_lifespan$lifespan))` days with `r round(mean(oarch_lifespan$commit_rate, na.rm = TRUE), 2)` commits per day.
<!-- TODO: turn all these percentages into code -->
Many projects are active for only a short period of time: about `r percent(p_lifespan_30days)` less than 30 days, `r percent(p_lifespan_90days)` less than 90 days, and `r percent(p_lifespan_365days)` less than a year.
However, the vast majority (all but `r english(n_1commit)`) do have more than one commit, suggesting that use of GitHub as a pure host for already-finished projects is not common; some degree of iteration, if not collaboration, is almost always present.
The longest-lived projects have been active for between `english(round(lifespan_top10_range / 365)[1])` and `english(round(lifespan_top10_range / 365)[2])` years.
The most active projects see up to `round(max_commit_rate)` commits per day, but the majority of repositories (`r percent(p_commit_rate_1)`) receive less one commit per day.

### 2023-10-12 analysis/openarchaeo_collaboration.qmd
Add statistics on lifespan and longevity

library("fs")
library("geomtextpath")
library("ggiraph")
library("ggraph")
library("gt")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("scales")
library("tidygraph")
library("tidyverse")
```

>>>>

library("english")
library("fs")
library("geomtextpath")
library("ggiraph")
library("ggraph")
library("gt")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("scales")
library("tidygraph")
library("tidyverse")
```
